### 2019-12-12 paper/6_Conclusions.tex

\section{Conclusions}
\noindent We have presented a method by which one can determine the level of non-determinism of a physics engine used in V\&V application of CAV simulations. The method consists of five stages, starting with designing the experiment, then suggesting the various settings that can improve or worsen the performance and finally focuses on running the experiment and analysing the output. \\\\
A case study is then used to show the implementation of the method introduced, where several tests were setup and run to explore the different interactions between the different agents in a simulator called CARLA. It was found that for the computer used (with specs of----) the level of non determinism is sufficiently low to assume the simulator to be deterministic as long as the CPU + GPU utilisation is below 50\% and that tests are terminated as soon as a collision of any type is detected. %since it was noticed that collisions cause the level of non determinism to jump remarkable invalidating the assumption of the simulator is deterministic.
% \TODO{Add the conclusion mentioned by Thales}