### 2021-04-12 document/ieee_trans/ieeetrans_template.tex

\documentclass[letterpaper, 10 pt, journal, twoside]{IEEEtran}
% \input{archive/main.config.tex}
\usepackage[maxnames=6,firstinits=true,doi=false,url=true,isbn=false]{biblatex}
% \addbibresource{IROS.bib}
\addbibresource{../source/cav_determinism.bib}
\usepackage{array}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}
\usepackage{threeparttable}
\usepackage{balance} %balance columns
\hyphenation{op-tical net-works semi-conduc-tor}
\usepackage{enumitem}
\usepackage{hyperref} % Add  [ocgcolorlinks,pdfusetitle] before hyperref for colored links
% \usepackage{subfig}
\usepackage{subcaption}
\usepackage{dblfloatfix}
\usepackage{booktabs}
\usepackage{authblk}
\newcommand{\TODO}[1]{{\color{red}\textbf{TODO: #1}}}
% \pagenumbering{gobble} %suppresses page numbers
\usepackage{balance} %balance columns
%\usepackage{todonotes}
\usepackage[disable]{todonotes}
\usepackage{siunitx}
% \usepackage{draftwatermark}
\usepackage{flushend}

%-------------------------------------------------------------------------------

\begin{document}

% ***************************************************
%  TITLE
% ***************************************************
\title{On Determinism of Game Engines used for Simulation-based Autonomous Vehicle Verification}
% \title{On the Precision of Game Engines used for Simulation-based Autonomous Vehicle Verification}
% \author[1,3]{Abanoub Ghobrial\thanks{$^{1}${\footnotesize \{abanoub.ghobrial, greg.chance, kevin.mcareavey, kerstin.eder\}@bristol.ac.uk}}}
% \author[1,3]{Greg Chance}
% \author[1,3]{Kevin McAreavey}
% \author[2,3]{Severin Lemaignan\thanks{$^{2}${\footnotesize \{severin.lemaignan, tony.pipe\}@brl.ac.uk}}}
% \author[2,3]{Tony Pipe}
% \author[1,3]{Kerstin Eder}
% \affil[1]{Trustworthy Systems Laboratory, University of Bristol, Bristol, UK}
% \affil[2]{University of West of England, Bristol, UK}
% \affil[3]{Bristol Robotics Laboratory, Bristol, UK}

% \author{Abanoub Ghobrial\thanks{$^{1}${\footnotesize \{abanoub.ghobrial, greg.chance, kevin.mcareavey, kerstin.eder\}@bristol.ac.uk}}}
% \author{Greg Chance}
% \author{Kevin McAreavey}
% \author{Severin Lemaignan\thanks{$^{2}${\footnotesize \{severin.lemaignan, tony.pipe\}@brl.ac.uk}}}
% \author{Tony Pipe}

\author{Greg Chance, Abanoub Ghobrial, Kevin McAreavey, S\'everin Lemaignan, Tony Pipe, Kerstin Eder 
\thanks{{\footnotesize
Manuscript  
received ...;
revised ...;  
accepted.... 
Date of publication ...;
date of current version ....

\textit{Greg Chance and Abanoub Ghobrial contributed equally to this paper and are both corresponding authors.}
%

This research has in part been funded by the ROBOPILOT and CAPRI projects. Both projects are part-funded by the Centre for Connected and Autonomous Vehicles (CCAV), delivered in partnership with Innovate UK under grant numbers 103703 (CAPRI) and 103288 (ROBOPILOT), respectively.
% The Associate Editor for this paper was ....

Greg Chance (e-mail: greg.chance@bristol.ac.uk), 
Abanoub Ghobrial (e-mail: abanoub.ghobrial@bristol.ac.uk), 
Kevin McAreavey (e-mail: kevin.mcareavey@bristol.ac.uk), 
and 
Kerstin Eder (e-mail: kerstin.eder@bristol.ac.uk) 
are with the Trustworthy Systems Lab, Department of Computer Science, University of Bristol, Merchant Ventures Building, Woodland Road,  Bristol, BS8 1UQ, United Kingdom.

S\'everin Lemaignan (e-mail: severin.lemaignan@brl.ac.uk)
and
Tony Pipe (e-mail: tony.pipe@brl.ac.uk), 
are with the Bristol Robotics Lab, T Block, University of the West of England,Frenchay, Coldharbour Ln, Bristol, BS34 8QZ, United Kingdom.

Digital Object Identifier ....
}}}
% \affil{Trustworthy Systems Laboratory, University of Bristol, Bristol, UK}
% \affil{University of West of England, Bristol, UK}
% \affil{Bristol Robotics Laboratory, Bristol, UK}
%
\markboth{IEEE TRANSACTIONS ON INTELLIGENT TRANSPORTATION SYSTEMS VOL. ..., NO. ..., date...}{Chance, Ghobrial \MakeLowercase{\textit{et al.}}: On Determinism of Game Engines used for Simulation-based Autonomous Vehicle Verification}
%
\maketitle

% ***************************************************
%  Abstract
% ***************************************************
\begin{abstract}
\noindent 
\input{../source/abstract.tex}
\end{abstract}

\begin{IEEEkeywords}
Autonomous Driving, Autonomous Vehicles, Determinism, Physics Engines,  Verification and Validation (V\&V), Simulation, Testing
\end{IEEEkeywords}
\IEEEpeerreviewmaketitle

% ***************************************************
%  Main Body
% ***************************************************
\input{../source/body.tex}
\textit{Greg Chance and Abanoub Ghobrial contributed equally to this paper.}

% ***************************************************
%  Bib
% ***************************************************
\printbibliography

% ***************************************************
%  Bio
% ***************************************************
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{../other/mugs/greg.png}}]{Greg Chance}
Greg Chance received the B.Sc. (Hons.) and Ph.D. degrees in physics from the University of Bath, Bath, U.K., in 2001 and 2005, respectively. He was previously at the Bristol Robotics Lab and is currently a Senior Research Associate with the Trustworthy Systems Laboratory, University of Bristol, Bristol, U.K. His current research interests are simulation-based verification for autonomous systems and cybersecurity. He has ten years industrial experience researching for Oxford Instruments, U.K., and BAE Systems, U.K. Dr Chance is a Chartered Engineer and a member of the Institute of Physics and holds a patent for a novel plasma control method. 
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{../other/mugs/abanoub}}]{Abanoub Ghobrial}
Abanoub Ghobrial received the MEng degree in mechanical engineering from the University of Bristol, Bristol,  U.K., in 2018. He is currently pursuing the Ph.D. degree in computer science at the University of Bristol and is a part-time Research Associate with the Trustworthy Systems Lab, Bristol,  U.K. From 2018 to 2020, he was a full-time Research Associate with the Trustworthy Systems Lab. His current research interests are techniques to allow self-managing of autonomous safety-critical systems via continual learning during operation; and the development of simulation-based verification techniques for autonomous systems.
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{../other/mugs/kevin.jpg}}]{Kevin McAreavey}
Kevin McAreavey received a BSc (Hons) in computer science from Queen's University Belfast (QUB) in 2010 followed by a PhD in AI in 2014. He is currently a Senior Research Associate with the Intelligent Systems Laboratory at the University of Bristol (UOB), working on explainable AI for AI-enabled domestic life. Previously he has worked with the Trustworthy Systems Laboratory at UOB and as a Research Fellow at QUB. His research interests are broadly in the areas of knowledge representation and reasoning, agents and multi-agent systems, and automated planning.
\end{IEEEbiography}

\begin{IEEEbiography}{Severin Lemaignan}
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{../other/mugs/tony}}]{Tony Pipe}
Tony Pipe received his BSc (Hons.) degree in Electronic Engineering from Warwick in 1979 and, after an eight-year period in industry that included running a small company, received his PhD in Robotics from the University of the West of England (UWE) in 1997. Since 2010 he has been a Professor of Robotics and Autonomous Systems at UWE. He was a founding member of the Bristol Robotics Laboratory (BRL), and acted as one of its two Deputy Directors from its inception in 2006 until 2020. His research interests focus on safe and correct operation whilst in close-proximity to human beings. He has co-authored over 200 international refereed publications, supervised 35 PhDs, and created a BRL income of >£10M since 2013. He has acted as BRL-UWE PI on the EPSRC grant ‘Trustworthy Robotic Assistants’, the Programme/Hub grants ‘Robotics for Nuclear Environments’ and ‘National Centre for Nuclear Robotics’, and the Centre for Doctoral Training in Robotics and Autonomous Systems. Prof. Pipe has been involved with Connected Autonomous Vehicle research and development since 2015 via the Innovate UK/CCAV funded projects VENTURER, FLOURISH, MultiCAV, CAPRI, ROBOPILOT and CAV-Forth.
\end{IEEEbiography}

\begin{IEEEbiography}{Kerstin Eder}
\end{IEEEbiography}

\end{document}